---
title: "R Notebook"
output: rmarkdown::github_document
---
Install swirl
```{r}
# install.packages("swirl")
packageVersion("swirl")
```

Load swirl
```{r}
library(swirl)
```

see a list of the variables in your workspace.
```{r}
ls()
```

clear your workspace

```{r}
rm(list=ls())
```

Start swirl
`swirl()`

What shall I call you? 

`Tim`

| Thanks, Tim. Let's cover a couple of quick housekeeping items before we begin  
| our first lesson. First of all, you should know that when you see '...', that  
| means you should press Enter when you are done reading and ready to continue.  

...  <-- That's your cue to press Enter to continue  

| Also, when you see 'ANSWER:', the R prompt (>), or when you are asked to  
| select from a list, that means it's your turn to enter a response, then press  
| Enter to continue.  

Select 1, 2, or 3 and press Enter   

1: Continue.  
2: Proceed.  
3: Let's get going!  

Selection:  

`1`

| You can exit swirl and return to the R prompt (>) at any time by pressing the  
| Esc key. If you are already at the prompt, type bye() to exit and save your  
| progress. When you exit properly, you'll see a short message letting you know  
| you've done so.  

| When you are at the R prompt (>):  
| -- Typing skip() allows you to skip the current question.  
| -- Typing play() lets you experiment with R on your own; swirl will ignore  
| what you do...  
| -- UNTIL you type nxt() which will regain swirl's attention.  
| -- Typing bye() causes swirl to exit. Your progress will be saved.  
| -- Typing main() returns you to swirl's main menu.  
| -- Typing info() displays these options again.  

| Let's get started!  

...  

| To begin, you must install a course. I can install a course for you from the  
| internet, or I can send you to a web page  
| (https://github.com/swirldev/swirl_courses) which will provide course options  
| and directions for installing courses yourself. (If you are not connected to  
| the internet, type 0 to exit.)  

1: R Programming: The basics of programming in R  
2: Regression Models: The basics of regression modeling in R  
3: Statistical Inference: The basics of statistical inference in R  
4: Exploratory Data Analysis: The basics of exploring data in R  
5: Don't install anything for me. I'll do it myself.  

Selection: 

`1`

|======================================================================| 100%

| Course installed successfully!
        
        
| Please choose a course, or type 0 to exit swirl.

1: R Programming
2: Take me to the swirl course repository!
        
Selection: 

`1`

| Please choose a lesson, or type 0 to return to course menu.

1: Basic Building Blocks      2: Workspace and Files     
3: Sequences of Numbers       4: Vectors                 
5: Missing Values             6: Subsetting Vectors      
7: Matrices and Data Frames   8: Logic                   
9: Functions                 10: lapply and sapply       
11: vapply and tapply         12: Looking at Data         
13: Simulation                14: Dates and Times         
15: Base Graphics             

Selection: 

`1`
