---
title: "Week 4"
output: rmarkdown::github_document
---

# Table of Contents

- **The str Function**

- **Simulation**
  - Generating Random Numbers
  - Simulating a Linear Model
  - Random Sampling
  
- **R Profiler**
  - Part 1
  - Part 2
  
- **Quiz**

- **Programming Assignment**

- **Key Takeaway Functions**

# The str Function  

`str()`: Compactly display the internal "structure" of an R object

- A diagnostic function and an alternative to `summary`
- It is especially well suited to compactly display the (abbreviated) contents of (possibly nested) lists
- Roughly one line per basic object

**What's in this object?**  

```{r}
str(str)
```
It is a function that takes any object  

**Function**
```{r}
str(lm)
str(ls)
```

**Vector**
```{r}
x <- rnorm(100, 2, 4)
summary(x)
str(x)
```

**Factor**
```{r}
f <- gl(40, 10)
f

str(f)
summary(f) # displays the number of elements in each level
```

**Data Frame**
```{r}
library(datasets)
head(airquality)
summary(airquality)
str(airquality)
```

**Matrix**
```{r}
m <- matrix(rnorm(100), 10, 10)
str(m) # Matrix is a two-dimensional array
m[, 1]
```

**List**
```{r}
s <- split(airquality, airquality$Month)
str(s)
```

# Simulation  

## Generating Random Numbers  

Functions for *probability distributions* in R  

- `rnorm`: generate "**r**andom **norm**al" variates with a given *mean* and *standard deviation*
- `dnorm`: evaluate the "**norm**al probability **d**ensity" (with a given mean/SD) at a point (or vector of points)
- `pnorm`: evaluate the "**cumulative** distribution function" for a **norm**al distribution
- `rpois`: generate **r**andom **pois**son variates with a given rate

*Probability distribution functions* usually have "four" functions associated with them.  
The functions are prefixed with a:  

- `d` for density
- `r` for random number generation
- `p` for cumulative distribution
- `q` for quantile function

Working with the *normal* distributions requires using these "four" functions
```{r}
# dnorm(x, mean = 0, sd = 1, log = FALSE)
# pnorm(q, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)
# qnorm(p, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)
# rnorm(n, mean = 0, sd = 1)
```

If <img src="https://render.githubusercontent.com/render/math?math=$\Phi$"> is the *cumulative distribution function* for a standard normal distribution, then `pnorm(q)` = <img src="https://render.githubusercontent.com/render/math?math=$\Phi(q)$"> and `qnorm(p)` = <img src="https://render.githubusercontent.com/render/math?math=$\Phi^{-1}(p)$">  

```{r}
x <- rnorm(10)
x

x <- rnorm(10, 20, 2)
x

summary(x)
```
Setting the random number seed with `set.seed` ensures "reproducibility"  
Always set the random number seed when conducting a simulation!  

#### Normal Random Variables
```{r}
set.seed(1)
rnorm(5)

set.seed(1)
rnorm(5)
rnorm(5)

set.seed(1)
rnorm(5)
rnorm(5)
rnorm(5)
```

#### Poisson Random Variables
```{r}
rpois(10, 1)
rpois(10, 2)
rpois(10, 29)
```

```{r}
# Cumulative distribution
ppois(2, 2) # Pr(x <= 2)
ppois(4, 2) # Pr(x <= 4)
ppois(6, 2) # Pr(x <= 6)
```

## Simulating a Linear Model 

Suppose we want to simulate from the following linear model  

<img src="https://render.githubusercontent.com/render/math?math=$y = \beta_0 + \beta_1 x + \epsilon$">

where <img src="https://render.githubusercontent.com/render/math?math=$\epsilon \sim \mathcal{N}(0, 2^{2})$">.  Assume <img src="https://render.githubusercontent.com/render/math?math=$x \sim \mathcal{N}(0, 1^{2}), \: \beta_0 = 0.5 \: and \: \beta_1 = 2$">.

```{r}
set.seed(20)
x <- rnorm(100)
e <- rnorm(n = 100, mean = 0, sd = 2)

y <- 0.5 + 2 * x + e
summary(y)

plot(x, y)
abline(lm(y ~ x), col = "blue")

model <- lm(y ~ x)
summary(model)
```

What if `x` is binary?

```{r}
set.seed(10)
x <- rbinom(n = 100, size = 1, prob = 0.5)
e <- rnorm(n = 100, mean = 0, sd = 2)

y <- 0.5 + 2 * x + e
summary(y)
plot(x, y)
```

Suppose we want to simulate from a *Poisson* model where  

<img src="https://render.githubusercontent.com/render/math?math=$Y \sim Poisson(\mu)$">  
<img src="https://render.githubusercontent.com/render/math?math=$log \: \mu = \beta_0 + \beta_1 x$">  
and <img src="https://render.githubusercontent.com/render/math?math=$\beta_0 = 0.5 \: and \: \beta_1 = 0.3$">.  
We need to use the `rpois` function for this

```{r}
set.seed(1)
x <- rnorm(100)
log.mu <- 0.5 + 0.3 * x

y <- rpois(100, exp(log.mu))
summary(y)
plot(x, y)
```

## Random Sampling  

The `sample()` function draws randomly from a specified set of (scalar) objects allowing you to sample from arbitrary distributions.
```{r}
set.seed(1)

sample(1:10, 4)
sample(1:10, 4)

sample(letters, 5)

sample(1:10) # permutation
sample(1:10)

sample(1:10, replace = TRUE) # Sample with replacement
```

# R Profiler  

- "Profiling" is a systematic way to "examine how much time" is spent in different parts of a program
- Useful when trying to "optimize" your code

## Part 1  

#### system.time()  

- Takes an arbitrary R expression as input (can be wrapped in curly braces) and returns the amount of time taken to evaluate the expression
- Computes the time (*in seconds*) needed to execute an expression
- Returns an object of class `proc_time`
  - **user time:** time charged to the CPU(s) for this expression (for the computer)
  - **elapsed time:** "wall clock" time (for myself)

```{r}
# Elapsed time > user time
system.time(readLines("http://www.jhsph.edu"))
# The chunk of the time is just spent waiting for the network, or for the data 
# to go over the network and to come back to your computer
```

```{r}
# Elapsed time < user time
hilbert <- function(n) {
        i <- 1:n
        1 / outer(i - 1, i, "+")
}

x <- hilbert(1000)
system.time(svd(x))
```

Timing longer expressions:
```{r}
system.time({
        n <- 1000
        r <- numeric(n)
        for (i in 1:n) {
                x <- rnorm(n)
                r[i] <- mean(x)
        }
})
```

## Part 2  

#### Rprof()  

- The `Rprof()` function starts the profiler in R
  - R must be complied with profiler support (but this is usually the case)
- The `summaryRprof()` function summarizes the output from `Rprof()` (otherwise it's not readable)
- DO NOT use `system.time()` and `Rprof()` together or you will be sad
- `Rprof()` keeps track of the "function call stack" at regularly sampled intervals and tabulates how much time is spent in each function
- Default sampling interval is 0.02 seconds
- NOTE: If your code runs very quickly, the profiler is not useful, but then you probably don't need it in that case

```{r}
# lm(y ~ x)

sample.interval=10000
```

- The `summaryRprof()` function tabulates the R profiler output and calculates how much time is spent in which function
- There are two methods for normalizing the data
- `by.total` divides the time spent in each function by the total run time
- `by.self` does the same but first subtracts out time spent in functions above in the call stack

#### by.total

```{r}
# $by.total
#                           total.time         total.pct     self.time        self.pct
# "lm"                            7.41           100.00           0.30            4.05
# "lm.fit"                        3.50            47.23           2.99           40.35
# "model.frame.default"           2.24            30.23           0.12            1.62
# "eval"                          2.24            30.23           0.00            0.00
# "model.frame"                   2.24            30.23           0.00            0.00
# "na.omit"                       1.54            20.78           0.24            3.24
# "na.omit.data.frame"            1.30            17.54           0.49            6.61
# "lapply"                        1.04            14.04           0.00            0.00
# "[.data.frame"                  1.03            13.90           0.79           10.66
# "["                             1.03            13.90           0.00            0.00
# "as.list.data.frame"            0.82            11.07           0.82           11.07
# "as.list"                       0.82            11.07           0.00           11.00
```

#### by.self  
Tells you how much time is being spent in a given function, but after subtracting out all of the other time spent in lower level functions that it calls.

```{r}
# $by.self
#                          self.time          elf.pct     total.time       total.pct
# "lm"                          2.99            40.35           3.50           47.23
# "as.list.data.frame"          0.82            11.07           0.82           11.07
# "[.data.frame"                0.79            10.66           1.03           13.90
# "structure"                   0.73             9.85           0.73            9.85
# "na.omit.data.frame"          0.49             6.61           1.30           17.54
# "list"                        0.46             6.21           0.46            6.21
# "lm"                          0.30             4.05           7.41          100.00
# "model.matrix.default"        0.27             3.64           0.79           10.66
# "na.omit"                     0.24             3.24           1.54           20.78
# "as.character"                0.18             2.43           0.18            2.43
# "model.frame.default"         0.12             1.62           2.24           30.23
# "anyDuplicated.default"       0.02             0.27           0.02            0.27
```

#### summaryRprof() output  
Summarizes the output of `Rprof()` and gives "percent of time" spent in each function (with two types of normalization)
```{r}
# $sample.interval
## [1] 0.02

# $sampling.time        # the total amount of time that the expression took to run
## [1] 7.41             # = elapse time
```

# Quiz  

## Question 1  
**What is produced at the end of this snippet of R code?**  
```{r}
set.seed(1)
rpois(n = 5, lambda = 2)
```


## Question 2  
**What R function can be used to generate standard Normal random variables?**
(Answer): `rnorm`  

## Question 3  
**When simulating data, why is using the `set.seed()` function important?**  
**Select all that apply.**  
(Answer):   
It ensure that the sequence of random numbers starts in a specific place and its therefore reproducible.  

## Question 4  
**Which function can be used to evaluate the inverse cumulative distribution function for the Poisson distribution?**  
(Answer): `qpois`  

## Question 5  
**What does the following code do?**   
```{r}
set.seed(10)
x <- rep(0:1, each = 5)
e <- rnorm(10, 0, 20)
y <- 0.5 + 2 * x + e
```

(Answer): Generate data from a Normal linear model

## Question 6  
**What R function can be used to generate Binomial random variables?**  
(Answer): `rbinom()`  

## Question 7  
**What aspect of the R runtime does the profiler keep track of when an R expression is evaulated?**  
(Answer): The function call stack  

## Question 8  
**Consider the following R code**  
```{r}
library(datasets)
Rprof()
# fit <- lm(y ~ x1 + x2)
Rprof(NULL)
```
**(Assume that y, x1, and x2 are present in the workspace).**  
**Without running the code, what percentage of the run time is spent in the `lm()` function, based on the `by.total` method of normalization shown in `summaryRprof()`?**  
(Answer): 100%

## Question 9  
**When using `system.time()`, what is the user time?**  
(Answer): It is the time spent by the CPU evaluating an expression  

## Question 10  
**If a computer has more than one available processor and R is able to take advantage of that, then which of the following is true when using `system.time()`?**  
(Answer): Elapsed time may be smaller than user time  


# Programming Assignment  

```{r}
getwd()
setwd("/Users/hyeonwooyang/Desktop/Desktop/5_Coursera/1_R_Programming/git_project/data")
```

```{r, setup, echo = FALSE}
knitr::opts_knit$set(root.dir = '/Users/hyeonwooyang/Desktop/Desktop/5_Coursera/1_R_Programming/git_project/data')
```

## 1. Plot the 30-day mortality rates for heart attack

```{r}
outcome <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
# head(outcome)
# ncol(outcome)
# names(outcome)
# summary(outcome)
str(outcome)
```

```{r}
outcome[, 11] <- as.numeric(outcome[, 11])
hist(outcome[, 11])
```

## 2. Finding the best hospital in a state  

```{r}
best <- function(state, outcome) {
        ## Read outcome data
        data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
        a <- split(data, data[, "State"])
        
        ## Check that state and outcome are valid
        if (!state %in% names(a)) {
                return(paste("Error in best(", state, ", ", outcome, ") : invalid state", sep = ""))
        }
        else if (!outcome %in% c("heart attack", "heart failure", "pneumonia")) {
                return(paste("Error in best(", state, ", ", outcome, ") : invalid outcome", sep = ""))
        }
        
        ## Return hospital name in that state with lowest 30-day death
        ## rate
        b <- a[[state]]
        
        x <- b[, "Hospital.Name"]
        
        if (outcome == "heart attack") {
                c <- 11
        }
        else if (outcome == "heart failure") {
                c <- 17
        }
        else {
                c <- 23
        }
        y <- as.numeric(b[, c])

        combine <- data.frame(x, y)

        good <- complete.cases(x, y)
        newdata <- combine[good, ]

        order <- newdata[order(newdata[, 2], newdata[, 1]), ]
        # = newdata[with(newdata, order(y, x)), ]

        ans <- unname(order[1, 1])
        
        return(ans)
}
```

```{r}
dump("best", file = "best.R")
```

```{r}
rm(list = ls())
source("best.R")
```

```{r warning = FALSE}
best("TX", "heart attack")
best("TX", "heart failure")
best("MD", "heart attack")
best("MD", "pneumonia")
best("BB", "heart attack")
best("NY", "hert attack")
```

## 3. Ranking hospitals by outcome in a state

```{r}
rankhospital <- function(state, outcome, num = "best") {
        ## Read outcome data
        data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
        a <- split(data, data[, "State"])
        
        ## Check that state and outcome are valid
        if (!state %in% names(a)) {
                return(paste("Error in best(", state, ", ", outcome, ") : invalid state", sep = ""))
                # stop(paste("Error in best(", state, ", ", outcome, ") : invalid state", sep = ""))
        }
        else if (!outcome %in% c("heart attack", "heart failure", "pneumonia")) {
                return(paste("Error in best(", state, ", ", outcome, ") : invalid outcome", sep = ""))
                # stop(paste("Error in best(", state, ", ", outcome, ") : invalid outcome", sep = ""))
        }
        
        ## Return hospital name in that state with the given rank
        ## 30-day death rate
        b <- a[[state]]
        
        x <- b[, "Hospital.Name"]
        
        if (outcome == "heart attack") {
                c <- 11
        }
        else if (outcome == "heart failure") {
                c <- 17
        }
        else {
                c <- 23
        }
        y <- as.numeric(b[, c])

        combine <- data.frame(x, y)
        colnames(combine) <- c("Hospital.Name", "Rate")

        good <- complete.cases(x, y)
        newdata <- combine[good, ]
        
        order <- newdata[order(newdata[, 2], newdata[, 1]), ]
        # = newdata[with(newdata, order(y, x)), ]

        order$Rank <- 1:nrow(order)
        
        if (num == "best") {
                ans <- unname(order[1, 1])
        }
        else if (num == "worst") {
                ans <- unname(order[nrow(order), 1])
        }
        else if (num > nrow(order)) {
                ans <- NA
        }
        else {
              ans <- unname(order[num, 1])  
        }
        return(ans)
}
```

```{r}
dump("rankhospital", file = "rankhospital.R")
```

```{r warning = FALSE}
source("rankhospital.R")
rankhospital("TX", "heart failure", 4)
rankhospital("MD", "heart attack", "worst")
rankhospital("MN", "heart attack", 5000)
```

## 4. Ranking hospitals in all states  
```{r}
rankall <- function(outcome, num = "best") {
        ## Read outcome data
        data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
        a <- split(data, data[, "State"])
        
        ## Check that outcome is valid
        if (!outcome %in% c("heart attack", "heart failure", "pneumonia")) {
                return(paste("Error in best(", state, ", ", outcome, ") : invalid outcome", sep = ""))
                # stop(paste("Error in best(", state, ", ", outcome, ") : invalid outcome", sep = ""))
        }
        
        ## For each state, find the hospital of the given rank
        if (outcome == "heart attack") {
                c <- 11
        }
        else if (outcome == "heart failure") {
                c <- 17
        }
        else {
                c <- 23
        }
                
        zList <- lapply(a, function(df) {
                        x <- df[, "Hospital.Name"]
                        y <- as.numeric(df[, c])
                
                        combine <- data.frame(x, y)
                        colnames(combine) <- c("Hospital.Name", "Rate")
                
                        good <- complete.cases(x, y)
                        newdata <- combine[good, ]
                
                        order <- newdata[order(newdata[, 2], newdata[, 1]), ]
                        # = newdata[with(newdata, order(y, x)), ]
                
                        order$Rank <- 1:nrow(order)
                
                        if (num == "best") {
                                ans <- unname(order[1, 1])
                        }
                        else if (num == "worst") {
                                ans <- unname(order[nrow(order), 1])
                        }
                        else if (num > nrow(order)) {
                                ans <- NA
                        }
                        else {
                                ans <- unname(order[num, 1])  
                        }
                        return(ans)
                })
        
        ## Return a data frame with the hospital names and the
        ## (abbreviated) state name
        df2 <- data.frame()
        
        for (i in seq_along(zList)) {
                z <- c(zList[[i]], names(zList[i]))
                df2 <- rbind(df2, z)
        }
        
        colnames(df2) <- c("hospital", "state")
        return(df2)
}
```

```{r}
dump("rankall", file = "rankall.R")
```

```{r warning = FALSE}
source("rankall.R")
head(rankall("heart attack", 20), 10)
tail(rankall("pneumonia", "worst"), 3)
tail(rankall("heart failure"), 10)
```

## Programming Assignment Quiz  

## Question 1
```{r warning = FALSE}
best("SC", "heart attack")
```

## Question 2
```{r warning = FALSE}
best("NY", "pneumonia")
```

## Question 3
```{r warning = FALSE}
best("AK", "pneumonia")
```

## Question 4
```{r warning = FALSE}
rankhospital("NC", "heart attack", "worst")
```

## Question 5
```{r warning = FALSE}
rankhospital("WA", "heart attack", 7)
```

## Question 6
```{r warning = FALSE}
rankhospital("TX", "pneumonia", 10)
```

## Question 7
```{r warning = FALSE}
rankhospital("NY", "heart attack", 7)
```

## Question 8
```{r warning = FALSE}
r <- rankall("heart attack", 4)
as.character(subset(r, state == "HI")$hospital)
```

## Question 9
```{r warning = FALSE}
r <- rankall("pneumonia", "worst")
as.character(subset(r, state == "NJ")$hospital)
```

## Question 10
```{r warning = FALSE}
r <- rankall("heart failure", 10)
as.character(subset(r, state == "NV")$hospital)
```

# Key Takeaway Functions

- **The str Function**
  - `str()`
  - `gl(levels, repeat)`

- **Simulation**
  - Generating Random Numbers
    - `<img src="https://render.githubusercontent.com/render/math?math=">`
    - `rnorm(n, mean = 0, sd = 1)`
    - `rnorm(), dnorm(), pnorm(), rpois()`
    - `binorm(n, size, prob)`
    - `set.seed()`
   
  - Simulating a Linear Model
    - `rbinorm()`
    - `plot(x, y); abline(lm(y ~ x), col = "blue")`
    - `summary(lm(y ~ x))`
  
  - Random Sampling
    - `sample()`
    
- **R Profiler**
  - `system.time()`
  - `Rprof(), summaryRprof(), by.total, by.self`
  - `sample.internal`

- **Programming Assignment**
  - `hist()`
  - `element %in% vector`
  - `vector[order(vector[, 1]), ]`
  - `vector[order(vector[, 2], vector[, 1]), ]`
  - `names(df)`
  - `unname(matrix[1, 1])`
  - `stop()`
  - `colnames(df)`
  - `df$new_column <-`
  - `for (i in seq_along(vector))`

<https://www.stat.cmu.edu/~cshalizi/rmarkdown/#math-in-r-markdown>  
<https://bookdown.org/yihui/rmarkdown/html-document.html#mathjax-equations>  
<https://gist.github.com/a-rodin/fef3f543412d6e1ec5b6cf55bf197d7b>  
<https://rpruim.github.io/s341/S19/from-class/MathinRmd.html>  
<https://chartio.com/resources/tutorials/how-to-sort-a-data-frame-by-multiple-columns-in-r/>  

  