---
title: "Week1"
output: rmarkdown::github_document
---
# Table of Contents
- **Background Material**
  - Working Directory
    
- **Week 1: Getting Started and R Nuts and Bolts** 
  - Objects, Numbers, Attributes, Vectors, Mixing Objects
  - List, Matrices, Factors, Missing Values, Data Frames, Names
  - Summary
  - Reading Data, Writing Data, Reading Large Tables
  - Connections: Interfaces to the Outside World
  - Subsetting: Basics, Lists, Matrices, Partial Matching, Removing Missing Values
  - Vectorized Operations
  
- **Quiz** 

- **Review of Attributes**

- **Key Takeaway Functions**

# Background Material

Get working directory
```{r}
getwd()
```

Set a new working directory
```{r}
setwd("/Users/hyeonwooyang/Desktop/Desktop/5_Coursera/1_R_Programming/git_project/data")
```
```{r, setup, echo = FALSE}
knitr::opts_knit$set(root.dir = '/Users/hyeonwooyang/Desktop/Desktop/5_Coursera/1_R_Programming/git_project/data')
```

Display a list of files in the working directory
```{r}
dir()
```

Create a function
```{r}
myfunction <- function() {
        x <- rnorm(100)
        mean(x)
}
```

Execute a function
```{r}
myfunction()
```

Display a list of objects in the current environment
```{r}
ls()
```

Source(bring) all functions (objects) in a different file:  
`source("mycode.R")`

# Week 1: Getting Started and R Nuts and Bolts 

Assignment operator, `<-`
```{r}
x <- 5
msg <- "hello"
```

Display the numeric object (vector), `x`, that has one element
```{r}
print(x)        # explicit printing
x               # auto-printing
```
*The [1] indicated that x is a `vector` and 5 is the first element*

Display the character object (vector), `msg`, that has one element
```{r}
print(msg)
msg
```

Create integer sequences
```{r}
x <- 1:20
x
```

## Objects
Everything in R is an `object`. `Atomic` classes of objects:  

- character  
- numeric (real numbers)  
- integer  
- complex  
- logical (TRUE/FALSE)  

The most basic object is a `vector` 

- A vector can only contain objects of the same class  
BUT: The one exception is a "list", which is represented as a vector but can contain objects of different classes

## Numbers
Infinity  
`Inf`

Not a Number (undefined value)  
`NaN`

## Attributes 
`name`, `dimnames`, `dimensions`, `class`, `length`   
function: `attributes()`

## Vectors
Create vectors of objects `c()`
```{r}
x <- c(0.5, 0.6)        # numeric
x <- c(TRUE, FALSE)     # logical
x <- c(T, F)            # logical
x <- c("a", "b", "c")   # character
x <- 9:29               # integer
x <- c(1+0i, 2+4i)      # complex
```

`vector()` function
```{r}
x <- vector("numeric", length = 10)
x
```

## Mixing Objects        
Coerces to the least common denominator
```{r}
y <- c(1.7, "a")        # character
y <- c(TRUE, 2)         # numeric
y <- c("a", TRUE)       # character
```

Explicit Coercion: `as.`
```{r}
x <- 0:6
x

class(x)

as.numeric(x)
as.logical(x)
as.character(x)
```

Nonsensical coercion results in NAs
```{r}
x <- c("a", "b", "c")

as.numeric(x)
as.logical(x)
as.complex(x)
```

## List
```{r}
x <- list(1, "a", TRUE, 1 + 4i)
x
```
*[[1]] Elements of a list*  
*[1] 1 Elements of a vector*

## Matrices
Matrices are `vectors` with a `dimension attribute`  
Must have every element be the same class (object)
```{r}
m <- matrix(nrow = 2, ncol = 3)
m

dim(m)
attributes(m)
```

Matrices are constructed `column-wise`
```{r}
m <- matrix(1:6, nrow = 2, ncol = 3)
m
```

Matrices can also be created directly from `vectors` by adding a `dimension attribute`
```{r}
m <- 1:10
m

dim(m) <- c(2, 5)
m

```

Another way to create matrices by `cbind()` and `rbind()`
```{r}
x <- 1:3
y <- 10:12

cbind(x, y)
rbind(x, y)
```

## Factors
Factors are used to represent `categorical data`  
Factors can be `un-ordered` or `ordered`  
Factors are vectors with `labels`  
```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"))
x

table(x)
unclass(x)
```

Setting the order of the levels
```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"),
            levels = c("yes", "no"))
x

table(x)
unclass(x)
```

## Missing Values
`NA`  
`NaN`  
`is.na()` is used to test objects if they are NA  
`is.nan()` is used to test for NaN  
NA values have a class also, so there are integer NA, character NA, etc.  
A NaN value is also NA but the converse is not true (NaN means Not a Number, i.e. undefined value)  
```{r}
is.na(NaN)
is.nan(NA)
```

## Data Frames
Data frames are used to store tabular data  
Represented as a special type of list where every element of the list has to have the same length  
Each element of the list: column, and the length of each element of the list: the number of rows  
However, each column doesn't have to be the same type  

`row.names` - every row has its own name  
Data frames are created by calling `read.table()` or `read.csv()`  
Can be converted to a matrix by calling `data.matrix()` - force objects will be coerced to be the same class

```{r}
x <- data.frame(foo = 1:4, bar = c(T, T, F, F))
x

nrow(x)
ncol(x)
```

## Names Attribute  
Very useful for writing readable code and self-describing objects
```{r}
x <- 1:3
names(x)

names(x) <- c("foo", "bar", "norf") # name for each element
x
names(x)

x <- list(a = 1, b = 2, c = 3)
x

m <- matrix(1:4, nrow = 2, ncol = 2)
dimnames(m) <- list(c("a", "b"), c("c", "d"))
m
```

## Summary
Data Types:  

- atomic classes: numeric, logical, character, integer, complex
- vectors, lists
- factors
- missing values
- data frames
- names

## Reading Data  
Tabular Data: `read.table()`, `read.csv()` (returns a dataframe in R)  
Lines of a text file: `readLines()`  
R Code files: `source()`  
R Code files: `dget()` for reading R objects that have been dparsed into text files  
Reading binary objects into R: `load()` for reading in saved workspaces, `unserialize()` for reading single R objects in binary form  

## Writing Data
`write.table()`  
`writeLines()`  
`dump()`  
`dput()`  
`save()`  
`serialize()`  

`read.table()` function  
```{r}
str(read.table)
```

- `file`, the name of a file, or a connection  
- `header`, logical indicating if the file has a header line  
- `sep`, a string indicating how the columns are separated  
- `colClasses`, a character vector indicating the class of each column in the dataset  
- `nrows`, the number of rows in the dataset  
- `comment.char`, a character string indicating the comment character  
- `skip`, the number of lines to skip from the beginning  
- `stringAsFactors`, should character variables be coded as factors? `defualt = TRUE(yes)`  

e.g. `data <- read.table("foo.txt")`  
`read.csv` = `read.table` except that the default separator is a comma  

## Reading Large Tables
To prevent R from choking:  
Read the "help" page for `read.table()`, which contains many hints  
Make a rough calculation of the memory required to store your dataset  
Set `comment.char = ""` if there are no commented lines in your file  

the `colClasses` argument  
initial <- `read.table("datatable.txt", nrows = 100)`  
classes <- `sapply(initial, class)`   
tabAll <- `read.table("datatable.text", colClasses = classes)`  

## Textual Data Formats  
`dumping` and `dputing` preserves metadata(sacrificing some readability), so that another user doesn't have to specify it all over again.  
advantage: editable  
disadvanage: not space-efficient  

`dput()`
```{r}
y <- data.frame(a = 1, b = "a")
dput(y)

dput(y, file = "y.R")
new.y <- dget("y.R")
new.y
```

`dput()` function, essentially writes R code, which can be used to reconstruct an R object  

`dump()`  
`dget()` can only be used on a single R object  
`dump()` can be used on multiple R objects  
```{r}
x <- "foo"
y <- data.frame(a = 1, b = "a")

dump(c("x", "y"), file = "data.R")
rm(x, y) # remove x and y
source("data.R")

y
x
```

## Connections: Interfaces to the Outside World
`file` opens a connection to a file  
`gzfile` opens a connection to a file compressed with gzip  
`bzfile` opens a connection to a file compressed with bzip2  
`url` opens a connection to a webpage  

```{r}
str(file) # "str()" displays the internal structure of a R object

```

- `description` is the name of the file    
- `open` is a code indicating    
- `r` read only  
- `w` writing (and initializina new file)  
- `a` appending  
- `rb`, `wb`, `ab` reading, writing, or appending in binary mode (Windows)  

`con <- file("foo.txt", "r")`  
`data <- read.csv(con)`  
`close(con)`  

is the same as  
`data <- read.csv("foo.txt")`  

```{r}
connection <- url("http://www.jhsph.edu", "r")
x <- readLines(connection)
head(x)
```

## Subsetting  
#### Basics    
`[` always returns an `object` of the same class as original (vector -> vector) (list -> list)  
Can be used to select more than one element of an object (there is one exception)  
`[[` extracts `elements` of a list or a dataframe  
Can only be used to extract a single element and the class of the returned object will not necessarily be a list or dataframe  
`$` extracts `elements` of a list or dataframe `by name`

```{r}
x <- c("a", "b", "c", "c", "d", "a")
x[1]
x[1:4]                          # numeric index
x[x > "a"]                      # logical index

u <- x > "a"
u
x[u]
```

#### Lists   
```{r}
x <- list(foo = 1:4, bar = 0.6)
x

x[1]

x[[1]]
x[[2]]

x$bar
x[["bar"]]      # = x[[2]]
x["bar"]        # = x[2]
```

```{r}
x <- list(foo = 1:4, bar = 0.6, baz = "hello")
x

x[c(1, 3)]

name <- "foo"
x[[name]]       # = x[[1]]
x$name          # element 'name' doesn't exist!
x$foo
```

```{r}
x <- list(a = list(10, 12, 14), b = c(3.14, 2.81))
x

x[[c(1, 3)]]
x[[1]][[3]]

x[[c(2, 1)]]
```

#### Matrices  
```{r}
x <- matrix(1:6, 2, 3)
x

x[1, 2]
x[2, 1]

x[1, ]
x[, 2]
```

By default, when a single element of a matrix is retrieved, it is returned as a `vector` of length 1 rather than a 1 x 1 matrix  
This behavior can be turned off by setting `drop = FALSE`  
```{r}
x[1, 2]
x[1, 2, drop = FALSE]

x[1, ]
x[1, , drop = FALSE]
```

#### Partial Matching      
```{r}
x <- list(aardvark = 1:5)
x

x$a
x[["a"]]
x [["a", exact = FALSE]]
```

#### Removing Missing Values      
```{r}
x <- c(1, 2, NA, 4, NA, 5)
is.na(x)
bad <- is.na(x)
x[!bad]
```

```{r}
x <- c(1, 2, NA, 4, NA, 5)
y <- c("a", "b", NA, "d", NA, NA)
complete.cases(x, y)

good <- complete.cases(x, y)
x[good]
y[good]
```

```{r}
airquality[1:6, ]
complete.cases(airquality[1:6, ])
good <- complete.cases(airquality)
airquality[good, ][1:6, ]
```

## Vectorized Operations
```{r}
x <- 1:4; y <- 6:9
x
y

x + y
x > 2
x >= 2
y == 8
x * y
x / y
```

```{r}
x <- matrix(1:4, 2, 2); y <- matrix(rep(10, 4), 2, 2)
x
y

x * y                   # element-wise multiplication
x / y
x %*% y                 # true matrix multiplication
```

# Week 1 Quiz
## Question 1
**The R language is a dialect of which of the following programming languages?**  
(Answer): S

## Question 2
**The definition of free software consists of four freedoms (freedoms 0 through 3).**  
**Which of the following is NOT one of the freedoms that are part of the definition? Select all that apply.**  
(Answer): all except the following  

- Freedom 0: to run the program, for any purpose  
- Freedom 1: to study how the program works, and adapt it to your needs  
- Freedom 2: to redistribute copies so you can help your neighbor  
- Freedom 3: improve the program, and release your improvements to the public, so that the whole community benefits  

## Question 3
**In R the following are all atomic data types EXCEPT: (Select all that apply)**  
(Answer):

- **Atomic data types:** numeric, complex, integer, logical, character
- **Data structures:** table, matrix, array, data frame, list

## Question 4
**If I execute the expression `x <- 4` in R,**  
**what is the class of the object `x` as determined by the `class()` function?**
```{r}
x <- 4
class(x)
```

## Question 5
**What is the class of the object defined by the expression `x <- c(4, "a", TRUE)`?**
```{r}
x <- c(4, "a", TRUE)
class(x)
```

## Question 6
**If I have two vectors `x <- c(1,3, 5)` and `y <- c(3, 2, 10)`,**  
**what is produced by the expression `rbind(x, y)`?**
```{r}
x <- c(1, 3, 5); y <- c(3, 2, 10)
rbind(x, y)
```

## Question 7
**A key property of vectors in R is that**  
(Answer): elements of a vector all must be of the same class

## Question 8
**Suppose I have a list defined as `x <- list(2, "a", "b", TRUE)`.**  
**What does `x[[2]]` give me?   Select all that apply**  
```{r}
x <- list(2, "a", "b", TRUE)
x[[2]]
```

## Question 9
**Suppose I have a vector `x <- 1:4` and a vector `y <- 2`.**  
**What is produced by the expression `x + y`?**
```{r}
x <- 1:4; y <- 2
x + y
```

## Question 10
**Suppose I have a vector `x <- c(17, 14, 4, 5, 13, 12, 10)` and I want to set all elements of this vector that are greater than 10 to be equal to 4.**  
**What R code achieves this? Select all that apply.**
```{r}
x <- c(17, 14, 4, 5, 13, 12, 10)

x[x > 10] <- 4
# OR
x[x >= 11] <- 4
```

## Question 11
**Use the Week 1 Quiz Data Set to answer questions 11-20.**  
**In the dataset provided for this Quiz, what are the column names of the dataset?**  
```{r}
getwd()
dir()
table <- read.csv("hw1_data.csv")

colnames(table)
```

## Question 12
**Extract the first 2 rows of the data frame and print them to the console.**  
**What does the output look like?**
```{r}
table[1:2, ]
head(table, 2)
```

## Question 13
**How many observations (i.e. rows) are in this data frame?**
```{r}
nrow(table)
```

## Question 14
**Extract the last 2 rows of the data frame and print them to the console.**  
**What does the output look like?**
```{r}
tail(table, 2)
```

## Question 15
**What is the value of `Ozone` in the 47th row?**
```{r}
table[47, "Ozone"]
```

## Question 16
**How many missing values are in the `Ozone` column of this data frame?**
```{r}
sum(is.na(table[, 1]))
```

## Question 17
**What is the mean of the `Ozone` column in this dataset?**  
**Exclude missing values (coded as NA) from this calculation.**

```{r}
mean(table[, 1], na.rm = TRUE)
# OR
x <- complete.cases(table[, 1])
mean(table[x, 1])
# OR
x <- is.na(table[, 1])
mean(table[!x, 1])
```

## Question 18
**Extract the subset of rows of the data frame where `Ozone` values are above 31 and `Temp` values are above 90.**  
**What is the mean of Solar.R in this subset?**
```{r}
x <- complete.cases(table[, "Ozone"], table[, "Temp"])
y <- table[x, ]
# View(y)

z <- y[y$Ozone > 31 & y$Temp > 90, ]
# View(z)
mean(z[, "Solar.R"])
```

## Question 19
**What is the mean of `Temp` when `Month` is equal to 6?**
```{r}
x <- table[table$Month == 6, ]
mean(x[, "Temp"])
```

## Question 20
**What was the maximum `ozone` value in the month of May (i.e. Month is equal to 5)?**
```{r}
x <- table[table$Month == 5, ]
max(x[, "Ozone"], na.rm = TRUE)
```

# Attributes (objects may have)

#### class  
*Vectors*  

- Atomic Vector: vectors consisted of only one class  
  - Data Types:  
    - **character:** `c("a", "swc")`  
    - **numeric** (real or decimal): `c(2, 15.5, Inf, NaN)`  
    - **integer:** `c(2L)` (the `L` tells R to store this as an integer    
    - **logical:** `c(TRUE, FALSE, NA)`  
    - **complex:** `c(1 + 4i)`  

  - **Factors**:  atomic vectors with "labels (= levels)", `factor(c("yes", "no"))`         
  - **Matrix** (two dimensional array):  atomic vectors, `matrix(1:6, nrow = 2, ncol = 3)` with "dimension" attributes  

- **List:**  vectors consisted of multiple classes, `list(1, "a")`    
  - **Data Frame:**  every element = same "length", but != same "type", `data.frame(foo = 1:2, bar = c(T, F))` with "dimension" attributes      

*Functions*  

*Date*  

Time   

- **`POSIXct`** (default)   
- **`POSIXlt`**


#### length

- **Number of elements in a *vector*:**
  - atomic vector (= factors, matrix)  

- **Number of elements in a *list*:**
  - list (= data frame)  
  
#### dim

- **NULL**:
  - atomic vector (= factors)
  - list
  
- **nrow**, **ncol:**
  - matrix (is an atomic vector with a "dimension" attribute)
  - data frame (can be converted to a matrix by calling `data.matrix()` - coerces objects to be the same class)
  
#### dimnames

- **NULL**
  - atomic vector (= factors)
  - list
  
- **list(rownames, colnames)**
  - matrix (if specified) `dimnames(m) <- list(c("row1", row2"), c("col1", "col2"))`
  - data frame

#### names

- **Name of each element in a vector or a list**
  - atomic vector (= factors, matrix) (if specified) `names(vector) <- c("one", "two")`
  - list (if specified) `list(foo = 1, bar = "a")`  
  - data frame

#### levels

- **NULL**
  - atomic vector (= matrix)
  - list (= data frame)
  
- **labels**
  - factors (is an atomic vector with "labels (= levels)")  

# Key Takeaway Functions
Comment out a block of code: `Cmd` + `Shift` + `C`  

- **Background Material**
  - Working Directory
    - *getwd(), dir(), ls(), rm(list=ls()), source()*
    
- **Week 1: Getting Started and R Nuts and Bolts** 
  - Numbers
    - *Inf, NaN*
  - Mixing Objects
    - *as.numeric(), as.logical(), as.character()*
  - Matrices
    - *m <- 1:10, dim(m) <- c(nrow, ncol)*
    - *matrix(rep(10, 4), 2, 2)*
    - *cbind(), rbind()*
  - Factors
    - *x <- factor(c("yes", "yes", "no"), levels = c("yes", "no"))*
  - Missing Values  
    - *is.na(), is.nan()*
    - *is.na(NaN) = TRUE, is.nan(NA) = FALSE*
  - Data Frames
    - *nrow(), ncol()*
    - *rownames(), colnames()*
  - Reading Data
    - *read.table(), read.csv(), readLines()*
  - Reading Large Tables
    - *dput(), dump(), dget()*  
  - Connections: Interfaces to the Outside World
    - *str()*
  - Subsetting: Removing Missing Values
    - *x[!is.na(x)], complete.cases(x, y)*
  
- **Quiz** 
  - *head(table, 2), tail(table, 2)* 
  - *table[, "Ozone"], table[table$Ozone > 5, ]* 
  - *max(vector, na.rm = TRUE)*
