---
title: "Week 3"
output: rmarkdown::github_document
---

# Table of Contents  

- **Loop Functions**  
  - lapply
  - sapply
  - apply
  - mapply
  - tapply
  - split

- **Debugging Tools**
  - Diagnosing the Problem
  - Basic Tools
  - Using the Tools
  - Summary
  
- **Quiz**

- **Programming Assignment**

- **Key Takeaway Functions**

# Loop Functions  

- `lapply`: Loop over a "list" and evaluate a function on each element
- `sapply`: Same as `lapply` but try to "simplify" the result
- `apply`: Apply a function over the "margins" of an "array"
- `mapply`: "Multivariate" version of `lapply`
- `tapply`: Apply a function over "subsets" of a "vector"

An auxiliary function `split` is also useful, particularly in conjunction with `lapply`. Splits objects into sub-pieces.

## lapply    
Takes three arguments  

1. a list `X` (if X is not a list, it will be coerced to a list using `as.list`)
2. a function `FUN`
3. `...`

```{r}
lapply
```

```{r}
x <- list(a = 1:5, b = rnorm(10))
lapply(x, mean)
```
`lapply` always returns a list.  

```{r}
x <- 1:4
lapply(x, runif, min = 0, max = 10)
```

`lapply` makes heavy use of *anonymous* functions, functions that have no names.
```{r}
x <- list(a = matrix(1:4, 2, 2), b = matrix(1:6, 3, 2))
x
```

An anonymous function for extracting the first column of each matrix.
```{r}
lapply(x, function(elt) {elt[, 1]})
```

## sapply
`sapply` will try to simplify the result of `lapply` if possible.

- If the result is a list where every element is "length 1", then "a vector" is returned
- If the result is a list where every element is a vector of "the same length" (> 1), "a matrix" is returned
- If it can't figure things out, "a list" is returned

```{r}
x <- list(a = 1:4, b = rnorm(10), c = rnorm(20, 1), d = rnorm(100, 5))
lapply(x, mean)
sapply(x, mean)
```

## apply   
`apply` is used to evaluate a function (often an anonymous one) over the margins of an **array**.

- It is most often used to apply a function to the rows or columns of a matrix
- It can be used with general arrays, e.g. taking the average of an array of matrices
- It is not really faster than writing a loop, but it works in one line!

```{r}
str(apply)
```

- `X` is an **array**
- `MARGIN` is an integer vector indicating which margins should be "retained".
- `FUN` is a function to be applied
- `...` is fr other arguments to be passed to `FUN`

```{r}
x <- matrix(rnorm(200), 20, 10) # dimension 1 has 20 rows, and dimension 2 has 10 columns
apply(x, 2, mean) 
apply(x, 1, sum)
```

- `rowSums` = `apply(x, 1, sum)`
- `rowMeans` = `apply(x, 1, mean)`
- `colSums` = `apply(x, 2, sum)`
- `colMeans` = `apply(x, 2, mean)`

```{r}
x <- matrix(rnorm(200), 20, 10)
apply(x, 1, quantile, probs = c(0.25, 0.75))
```

```{r}
a <- array(rnorm(2 * 2 * 10), c(2, 2, 10))
apply(a, c(1, 2), mean) # collapse (average over) the 3rd dimension
apply(a, 1, mean)
apply(a, 2, mean)
```

## mapply    
`mapply` is a multivariate apply of sorts which applies a function in parallel over a set of arguments.  
`mapply` can take multiple list arguments and then apply a function to the elements of thos of the multipe lists in parallel.

```{r}
str(mapply)
```

- `FUN` is a function to apply (the number of arguments of `FUN` >= the number of lists to be passed)
- `...` contains arguments to apply over
- `MoreArgs` is a list of other arguments to `FUN`.
- `Simplify` indicates whether the result should be simplified

The following is tedious to type
```{r}
list(rep(1, 4), rep(2, 3), rep(3, 2), rep(4, 1))
```
Instead we can do
```{r}
mapply(rep, 1:4, 4:1)
```

## tapply    
`tapply` is used to apply a function over subsets of a vector.
```{r}
str(tapply)
```

- `X` is a vector
- `INDEX` is a factor or a list of factors (or else they are coerced to factors) "same length as `X`"
- `FUN` is a function to be applied
- `...` contains other arguments to be passed `FUN`
- `simplify`, should we simplify the result?

```{r}
x <- c(rnorm(10), runif(10), rnorm(10, 1))
x

f <- gl(3, 10)
f

tapply(x, f, mean)
tapply(x, f, mean, simplify = FALSE)
```

## split  
Always returns a list.  
`tapply` without summary statistics  
`split` takes a vector or other objects and splits it into groups determined by a factor or a list of factors.
```{r}
str(split)
```

- `x` is a vector (or list) or data frame
- `f` is a factor (or coerced to one) or a list of factors
- `drop` indicates whether empty factors levels should be dropped

```{r}
x <- c(rnorm(10), runif(10), rnorm(10, 1))
f <- gl(3, 10)
split(x, f)
```

```{r}
lapply(split(x, f), mean) # = tapply(x, f, mean)
```

```{r}
library(datasets)
head(airquality)

# Split the data frame to "monthly" pieces
s <- split(airquality, airquality$Month)
airquality$Month

lapply(s, function(x) {colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE)})
sapply(s, function(x) {colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE)})
```

#### Splitting on More than One Level  
```{r}
x <- rnorm(10)
f1 <- gl(2, 5)
f2 <- gl(5, 2)

f1
f2

interaction(f1, f2)
```

Interactions can create empty levels.
```{r}
str(split(x, list(f1, f2))) # drop = FALSE
```

# Debugging Tools  

## Diagnosing the Problem  
Indications that something's not right  

- `message`: A generic notification/diagnostic message. Execution of the function continues
- `warning`: Something wrong, but not fatal. Execution of the function continues
- `error`: A fatal problem has occurred. Execution stops
- `condition`: Something unexpected can occur. Programmers can create their own conditions

#### Warning
```{r}
log(-1)
```

#### Error
```{r}
printmessage <- function(x) {
        if(x > 0) {
                print("x is greater than zero")
        }
        else {
                print("x is less than or equal to zero")
        }
        invisible(x) # prevents auto printing
}

printmessage(1)
# printmessage(NA)
## Error in if (x > 0) { : missing value where TRUE/FALSE needed
```

Fix this problem 
```{r}
printmessage2 <- function(x) {
        if(is.na(x)) {
                print("x is a missing value!")
        }
        else if (x > 0) {
                print("x is greater than zero")
        }
        else {
                print("x is less than or equal to zero")
        }
        invisible(x)
}

x <- log(-1)
printmessage2(x)
```

## Basic Tools  
The primary tools for debugging functions in R are  

- `traceback`: prints out the "function call stack" after an error occurs; tells you how many function calls you are in and where the error occurred 
- `debug`: flags a function for "debug" mode
- `browser`: suspends the execution of a function wherever it is called and puts the function in debug mode
- `trace`: allows you to insert debugging code into a function in specific places
- `recover`: "a error handler function". allows you to modify the error behavior, to browse the function call stack

## Using the Tools  
#### traceback  
```{r}
rm(list = ls())

# mean(x)
## Error in mean(x) : object 'x' not found

# traceback()
## 1: mean(x)

# lm(y ~ x)
## Error in eval(expr, envir, enclos) : object `y` not found

## traceback()
## 7: eval(expr, envir, enclos)         <- this is where the error occurred
## 6: eval(predvars, data, env)
## 5: model.frame.default(formula = y ~ x, drop.unused.levels = TRUE)
## 4: model.frame(formula = y ~ x, drop.unused.levels = TRUE)
## 3: eval(expr, envir, enclose)
## 2: eval(mf, parent.frame())
## 1: lm(y ~ x)
```

#### debug  
```{r}
# debug(lm)
# lm(y ~x)
## debugging in: lm(y ~ x)
## debug: {
##      ret.x <- x
##      ret.y <- y
##      cl <- match.call()
##      ...
##      if (!qr)
##              z$qr <- NULL
##      z
## }
## Browse[2]>                   <- This is the prompt. Now you are in the browser
```

```{r}
# Browse[2]> n                   <- 'n' stands for 'next'
## debug: ret.x <- x
# Browse[2]> n
## debug: ret.y <- y
# Browse[2]> n
## debug: cl <- match.call()
# Browse[2]> n
## debug: mf <- match.call(expand.dots = FALSE)
# Browse[2]> n
## debug: m <- match(c("formula", "data", "subset", "weights", "na.action", 
##      "offset"), names(mf), 0L)
```

#### recover   
```{r}
options(error = recover)
# read.csv("nosuchfile")
## Error in file(file, "rt") : cannot open the connection
## In addition: Warning message:
## In file(file, "rt") :
##      cannot open file 'nosuchfile': No such file or directory
##
## Enter a frame number, or 0 to exit
##
## 1: read.csv("nosuchfile")
## 2: read.table(file = file, header = header, sep = sep, quote = quote, dec =
## 3: file(file, "rt")          <- this is where the error occurred
##
## Selection:
```

## Summary  

- There are "three" main indications of a problem/condition: `message, warning, error`  
  - only an `error` is fatal
- When analyzing a function with a problem, make sure you can "reproduce" the problem, clearly state your "expectations" and how the output "differs" from your expectation
- Interactive debugging tools `traceback, debug, browser, trace,` and `recover` can be used to find problematic code in functions
- Debugging tools are not a substitute for thinking!

# Week 3 Quiz  

## Question 1  
**Take a look at the 'iris' dataset that comes with R.**  
*The data can be loaded with the code:**
```{r}
library(datasets)
data(iris)
```

**A description of the dataset can be found by running**
```{r}
?iris
head(iris)
```

**There will be an obejct called 'iris' in your workspace. In this dataset, what is the mean of 'Sepal.Length' for the species *virginica*?**  

**Please round your answer to the nearest whole number.**  
```{r}
x <- iris[iris$Species == "virginica", ]
mean(x[, "Sepal.Length"])

y <- tapply(iris$Sepal.Length, iris$Species, mean)
y

mean(y["virginica"])
```

(Answer): 7

## Question 2  
**Continuing with the 'iris' dataset from the previous Question, what R code returns a vector of the means of the variables 'Sepal.Length', 'Sepal.Width', and 'Petal.Width'?**  

```{r}
head(iris)

apply(iris[, 1:4], 2, mean)
```

## Question 3  
**Load the 'mtcars' dataset in R with the following code**  
```{r}
library(datasets)
data(mtcars)
```

**There will be an object names 'mtcars' in your workspace.**  
**You can find some information about the dataset by running**  
```{r}
?mtcars
head(mtcars)
```

**How can one calculate the average miles per gallon (mpg) by number of cylinders in the car (cyl)?**  
**Select all that apply.**  

```{r}
sapply(split(mtcars$mpg, mtcars$cyl), mean)

tapply(mtcars$mpg, mtcars$cyl, mean)

with(mtcars, tapply(mpg, cyl, mean))
```

## Question 4  
**Continuing with the 'mtcars' dataset from the previous Question,** 
**what is the absolute difference between the average horsepower of 4-cylinder cars and the average horsepower of 8-cylinder cars?**  

**(Please round your final answer to the nearest whole number)**  
```{r}
t <- tapply(mtcars$hp, mtcars$cyl, mean, simplify = FALSE)
t

abs(t[["4"]] - t[["8"]])
```

(Answer): 127

## Question 5  
**If you run**  
```{r}
debug(ls)
```
**What happens when you next call the 'ls()' function?**  
```{r}
# ls()
## debugging in: ls()
## debug: {
##     if (!missing(name)) {
##         pos <- tryCatch(name, error = function(e) e)
##         if (inherits(pos, "error")) {
##             name <- substitute(name)
##             if (!is.character(name)) 
##                 name <- deparse(name)
##             warning(gettextf("%s converted to character string", 
##                 sQuote(name)), domain = NA)
##             pos <- name
##         }
##     }
##     all.names <- .Internal(ls(envir, all.names, sorted))
##     if (!missing(pattern)) {
##         if ((ll <- length(grep("[", pattern, fixed = TRUE))) && 
##             ll != length(grep("]", pattern, fixed = TRUE))) {
##             if (pattern == "[") {
##                 pattern <- "\\["
##                 warning("replaced regular expression pattern '[' by  '\\\\['")
##             }
##             else if (length(grep("[^\\\\]\\[<-", pattern))) {
##                 pattern <- sub("\\[<-", "\\\\\\[<-", pattern)
##                 warning("replaced '[<-' by '\\\\[<-' in regular expression pattern")
##             }
##         }
##         grep(pattern, all.names, value = TRUE)
##     }
##     else all.names
## }
## Browse[2]> 
```

(Answer): Execution of 'ls' will suspend at the beginning of the function and you will be in the browser  

# Programming Assignment  



# Key Takeaway Functions

- **Loop Functions**  
  - lapply
    - *rnorm(n, mean = 0, sd = 1), runif(n, min = 0, max = 1)*
    
- **Debugging Tools**
  - Diagnosing the Problem
    - *invisible()*