---
title: "Week 3"
output: rmarkdown::github_document
---

# Table of Contents  

- **Loop Functions**  
  - lapply
  - sapply
  - apply
  - mapply
  - tapply
  - split

- **Debugging Tools**
  - Diagnosing the Problem
  - Basic Tools
  - Using the Tools
  - Summary
  
- **Quiz**

- **Programming Assignment**

- **Key Takeaway Functions**

# Loop Functions  

- `lapply`: Loop over a "list" and evaluate a function on each element
- `sapply`: Same as `lapply` but try to "simplify" the result
- `apply`: Apply a function over the "margins" of an "array"
- `mapply`: "Multivariate" version of `lapply`
- `tapply`: Apply a function over "subsets" of a "vector"

An auxiliary function `split` is also useful, particularly in conjunction with `lapply`. Splits objects into sub-pieces.

## lapply    
Takes three arguments  

1. a list `X` (if X is not a list, it will be coerced to a list using `as.list`)
2. a function `FUN`
3. `...`

```{r}
lapply
```

```{r}
x <- list(a = 1:5, b = rnorm(10))
lapply(x, mean)
```
`lapply` always returns a list.  

```{r}
x <- 1:4
lapply(x, runif, min = 0, max = 10)
```

`lapply` makes heavy use of *anonymous* functions, functions that have no names.
```{r}
x <- list(a = matrix(1:4, 2, 2), b = matrix(1:6, 3, 2))
x
```

An anonymous function for extracting the first column of each matrix.
```{r}
lapply(x, function(elt) {elt[, 1]})
```

## sapply
`sapply` will try to simplify the result of `lapply` if possible.

- If the result is a list where every element is "length 1", then "a vector" is returned
- If the result is a list where every element is a vector of "the same length" (> 1), "a matrix" is returned
- If it can't figure things out, "a list" is returned

```{r}
x <- list(a = 1:4, b = rnorm(10), c = rnorm(20, 1), d = rnorm(100, 5))
lapply(x, mean)
sapply(x, mean)
```

## apply   
`apply` is used to evaluate a function (often an anonymous one) over the margins of an **array**.

- It is most often used to apply a function to the rows or columns of a matrix
- It can be used with general arrays, e.g. taking the average of an array of matrices
- It is not really faster than writing a loop, but it works in one line!

```{r}
str(apply)
```

- `X` is an **array**
- `MARGIN` is an integer vector indicating which margins should be "retained".
- `FUN` is a function to be applied
- `...` is fr other arguments to be passed to `FUN`

```{r}
x <- matrix(rnorm(200), 20, 10) # dimension 1 has 20 rows, and dimension 2 has 10 columns
apply(x, 2, mean) 
apply(x, 1, sum)
```

- `rowSums` = `apply(x, 1, sum)`
- `rowMeans` = `apply(x, 1, mean)`
- `colSums` = `apply(x, 2, sum)`
- `colMeans` = `apply(x, 2, mean)`

```{r}
x <- matrix(rnorm(200), 20, 10)
apply(x, 1, quantile, probs = c(0.25, 0.75))
```

```{r}
a <- array(rnorm(2 * 2 * 10), c(2, 2, 10))
apply(a, c(1, 2), mean) # collapse (average over) the 3rd dimension
apply(a, 1, mean)
apply(a, 2, mean)
```

## mapply    
`mapply` is a multivariate apply of sorts which applies a function in parallel over a set of arguments.  
`mapply` can take multiple list arguments and then apply a function to the elements of thos of the multipe lists in parallel.

```{r}
str(mapply)
```

- `FUN` is a function to apply (the number of arguments of `FUN` >= the number of lists to be passed)
- `...` contains arguments to apply over
- `MoreArgs` is a list of other arguments to `FUN`.
- `Simplify` indicates whether the result should be simplified

The following is tedious to type
```{r}
list(rep(1, 4), rep(2, 3), rep(3, 2), rep(4, 1))
```
Instead we can do
```{r}
mapply(rep, 1:4, 4:1)
```

## tapply    
`tapply` is used to apply a function over subsets of a vector.
```{r}
str(tapply)
```

- `X` is a vector
- `INDEX` is a factor or a list of factors (or else they are coerced to factors) "same length as `X`"
- `FUN` is a function to be applied
- `...` contains other arguments to be passed `FUN`
- `simplify`, should we simplify the result?

```{r}
x <- c(rnorm(10), runif(10), rnorm(10, 1))
x

f <- gl(3, 10)
f

tapply(x, f, mean)
tapply(x, f, mean, simplify = FALSE)
```

## split  
Always returns a list.  
`tapply` without summary statistics  
`split` takes a vector or other objects and splits it into groups determined by a factor or a list of factors.
```{r}
str(split)
```

- `x` is a vector (or list) or data frame
- `f` is a factor (or coerced to one) or a list of factors
- `drop` indicates whether empty factors levels should be dropped

```{r}
x <- c(rnorm(10), runif(10), rnorm(10, 1))
f <- gl(3, 10)
split(x, f)
```

```{r}
lapply(split(x, f), mean) # = tapply(x, f, mean)
```

```{r}
library(datasets)
head(airquality)

# Split the data frame to "monthly" pieces
s <- split(airquality, airquality$Month)
airquality$Month

lapply(s, function(x) {colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE)})
sapply(s, function(x) {colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE)})
```

#### Splitting on More than One Level  
```{r}
x <- rnorm(10)
f1 <- gl(2, 5)
f2 <- gl(5, 2)

f1
f2

interaction(f1, f2)
```

Interactions can create empty levels.
```{r}
str(split(x, list(f1, f2))) # drop = FALSE
```

# Debugging Tools  

## Diagnosing the Problem  
Indications that something's not right  

- `message`: A generic notification/diagnostic message. Execution of the function continues
- `warning`: Something wrong, but not fatal. Execution of the function continues
- `error`: A fatal problem has occurred. Execution stops
- `condition`: Something unexpected can occur. Programmers can create their own conditions

#### Warning
```{r}
log(-1)
```

#### Error
```{r}
printmessage <- function(x) {
        if(x > 0) {
                print("x is greater than zero")
        }
        else {
                print("x is less than or equal to zero")
        }
        invisible(x) # prevents auto printing
}

printmessage(1)
# printmessage(NA)
## Error in if (x > 0) { : missing value where TRUE/FALSE needed
```

Fix this problem 
```{r}
printmessage2 <- function(x) {
        if(is.na(x)) {
                print("x is a missing value!")
        }
        else if (x > 0) {
                print("x is greater than zero")
        }
        else {
                print("x is less than or equal to zero")
        }
        invisible(x)
}

x <- log(-1)
printmessage2(x)
```

## Basic Tools  
The primary tools for debugging functions in R are  

- `traceback`: prints out the "function call stack" after an error occurs; tells you how many function calls you are in and where the error occurred 
- `debug`: flags a function for "debug" mode
- `browser`: suspends the execution of a function wherever it is called and puts the function in debug mode
- `trace`: allows you to insert debugging code into a function in specific places
- `recover`: "a error handler function". allows you to modify the error behavior, to browse the function call stack

## Using the Tools  
#### traceback  
```{r}
rm(list = ls())

# mean(x)
## Error in mean(x) : object 'x' not found

# traceback()
## 1: mean(x)

# lm(y ~ x)
## Error in eval(expr, envir, enclos) : object `y` not found

## traceback()
## 7: eval(expr, envir, enclos)         <- this is where the error occurred
## 6: eval(predvars, data, env)
## 5: model.frame.default(formula = y ~ x, drop.unused.levels = TRUE)
## 4: model.frame(formula = y ~ x, drop.unused.levels = TRUE)
## 3: eval(expr, envir, enclose)
## 2: eval(mf, parent.frame())
## 1: lm(y ~ x)
```

#### debug  
```{r}
# debug(lm)
# lm(y ~x)
## debugging in: lm(y ~ x)
## debug: {
##      ret.x <- x
##      ret.y <- y
##      cl <- match.call()
##      ...
##      if (!qr)
##              z$qr <- NULL
##      z
## }
## Browse[2]>                   <- This is the prompt. Now you are in the browser
```

```{r}
# Browse[2]> n                   <- 'n' stands for 'next'
## debug: ret.x <- x
# Browse[2]> n
## debug: ret.y <- y
# Browse[2]> n
## debug: cl <- match.call()
# Browse[2]> n
## debug: mf <- match.call(expand.dots = FALSE)
# Browse[2]> n
## debug: m <- match(c("formula", "data", "subset", "weights", "na.action", 
##      "offset"), names(mf), 0L)
```

#### recover   
```{r}
options(error = recover)
# read.csv("nosuchfile")
## Error in file(file, "rt") : cannot open the connection
## In addition: Warning message:
## In file(file, "rt") :
##      cannot open file 'nosuchfile': No such file or directory
##
## Enter a frame number, or 0 to exit
##
## 1: read.csv("nosuchfile")
## 2: read.table(file = file, header = header, sep = sep, quote = quote, dec =
## 3: file(file, "rt")          <- this is where the error occurred
##
## Selection:
```

## Summary  

- There are "three" main indications of a problem/condition: `message, warning, error`  
  - only an `error` is fatal
- When analyzing a function with a problem, make sure you can "reproduce" the problem, clearly state your "expectations" and how the output "differs" from your expectation
- Interactive debugging tools `traceback, debug, browser, trace,` and `recover` can be used to find problematic code in functions
- Debugging tools are not a substitute for thinking!

# Week 3 Quiz  

## Question 1  
**Take a look at the 'iris' dataset that comes with R.**  
*The data can be loaded with the code:**
```{r}
library(datasets)
data(iris)
```

**A description of the dataset can be found by running**
```{r}
?iris
head(iris)
```

**There will be an obejct called 'iris' in your workspace. In this dataset, what is the mean of 'Sepal.Length' for the species *virginica*?**  

**Please round your answer to the nearest whole number.**  
```{r}
x <- iris[iris$Species == "virginica", ]
mean(x[, "Sepal.Length"])

y <- tapply(iris$Sepal.Length, iris$Species, mean)
y

mean(y["virginica"])
```

(Answer): 7

## Question 2  
**Continuing with the 'iris' dataset from the previous Question, what R code returns a vector of the means of the variables 'Sepal.Length', 'Sepal.Width', and 'Petal.Width'?**  

```{r}
head(iris)

apply(iris[, 1:4], 2, mean)
```

## Question 3  
**Load the 'mtcars' dataset in R with the following code**  
```{r}
library(datasets)
data(mtcars)
```

**There will be an object names 'mtcars' in your workspace.**  
**You can find some information about the dataset by running**  
```{r}
?mtcars
head(mtcars)
```

**How can one calculate the average miles per gallon (mpg) by number of cylinders in the car (cyl)?**  
**Select all that apply.**  

```{r}
sapply(split(mtcars$mpg, mtcars$cyl), mean)

tapply(mtcars$mpg, mtcars$cyl, mean)

with(mtcars, tapply(mpg, cyl, mean))
```

## Question 4  
**Continuing with the 'mtcars' dataset from the previous Question,** 
**what is the absolute difference between the average horsepower of 4-cylinder cars and the average horsepower of 8-cylinder cars?**  

**(Please round your final answer to the nearest whole number)**  
```{r}
t <- tapply(mtcars$hp, mtcars$cyl, mean, simplify = FALSE)
t

abs(t[["4"]] - t[["8"]])
```

(Answer): 127

## Question 5  
**If you run**  
```{r}
debug(ls)
```
**What happens when you next call the 'ls()' function?**  
```{r}
# ls()
## debugging in: ls()
## debug: {
##     if (!missing(name)) {
##         pos <- tryCatch(name, error = function(e) e)
##         if (inherits(pos, "error")) {
##             name <- substitute(name)
##             if (!is.character(name)) 
##                 name <- deparse(name)
##             warning(gettextf("%s converted to character string", 
##                 sQuote(name)), domain = NA)
##             pos <- name
##         }
##     }
##     all.names <- .Internal(ls(envir, all.names, sorted))
##     if (!missing(pattern)) {
##         if ((ll <- length(grep("[", pattern, fixed = TRUE))) && 
##             ll != length(grep("]", pattern, fixed = TRUE))) {
##             if (pattern == "[") {
##                 pattern <- "\\["
##                 warning("replaced regular expression pattern '[' by  '\\\\['")
##             }
##             else if (length(grep("[^\\\\]\\[<-", pattern))) {
##                 pattern <- sub("\\[<-", "\\\\\\[<-", pattern)
##                 warning("replaced '[<-' by '\\\\[<-' in regular expression pattern")
##             }
##         }
##         grep(pattern, all.names, value = TRUE)
##     }
##     else all.names
## }
## Browse[2]> 
```

(Answer): Execution of 'ls' will suspend at the beginning of the function and you will be in the browser  

```{r}
undebug(ls)
```

# Assignment 2

## Example: Caching the Mean of a Vector  

In this example we introduce the `<<-` operator which can be used to assign a value to an object in an environment that is different from the current environment. Below are two functions that are used to create a special object that stores a numeric vector and cache's its mean.  

The first function, `makeVector` creates a special "vector", which is really a list containing a function to  

1. set the value of the vector  
2. get the value of the vector  
3. set the value of the mean  
4. get the value of the mean  

```{r}
rm(list=ls())
```

```{r}
makeVector <- function(x = numeric()) {
        m <- NULL
        set <- function(y) {
                x <<- y
                m <<- NULL
        }
        get <- function() x
        setmean <- function(mean) m <<- mean
        getmean <- function() m
        list(set = set, get = get,
             setmean = setmean,
             getmean = getmean)
}
```

```{r}
a <- makeVector(c(1, 2, 3, 4))
a
```

```{r}
ls(environment(makeVector))

a[["set"]]
```

#### set
```{r}
ls(environment(a[["set"]]))
get("x", environment(a[["set"]]))
get("m", environment(a[["set"]]))
```

#### get
```{r}
ls(environment(a[["get"]]))
get("x", environment(a[["get"]]))
get("m", environment(a[["get"]]))
```

#### setmean
```{r}
ls(environment(a[["setmean"]]))
get("x", environment(a[["setmean"]]))
get("m", environment(a[["setmean"]]))
```

#### getmean
```{r}
ls(environment(a[["getmean"]]))
get("x", environment(a[["getmean"]]))
get("m", environment(a[["getmean"]]))
```

The following function calculates the mean of the special "vector" created with the above function. However, it first checks to see if the mean has already been calculated. If so, it gets the mean from the cache and skips the computation. Otherwise, it calculates the mean of the data and sets the value of the mean in the cache via the `setmean` function.

```{r}
cachemean <- function(x, ...) {
        m <- x$getmean()
        if(!is.null(m)) {
                message("getting cached data")
                return(m)
        }
        data <- x$get()
        m <- mean(data, ...)
        x$setmean(m)
        m
}
```

```{r}
cachemean(a)

ls(environment(a[["set"]]))
get("x", environment(a[["set"]]))
get("m", environment(a[["set"]]))

get("x", environment(a[["get"]]))
get("m", environment(a[["get"]]))

get("x", environment(a[["setmean"]]))
get("m", environment(a[["setmean"]]))

get("x", environment(a[["getmean"]]))
get("m", environment(a[["getmean"]]))
```

```{r}
cachemean(a)
```

## Assignment: Caching the Inverse of a Matrix  

Matrix inversion is usually a costly computation and there may be some benefit to caching the inverse of a matrix rather than compute it repeatedly (there are also alternatives to matrix inversion that we will not discuss here). Your assignment is to write a pair of functions that cache the inverse of a matrix.

Write the following functions:

1. `makeCacheMatrix`: This function creates a special "matrix" object that can cache its inverse.
2. `cacheSolve`: This function computes the inverse of the special "matrix" returned by `makeCacheMatrix` above. If the inverse has already been calculated (and the matrix has not changed), then the `cachesolve` should retrieve the inverse from the cache.

Computing the inverse of a square matrix can be done with the `solve` function in R. For example, if X is a square invertible matrix, then `solve(X)` returns its inverse.

For this assignment, assume that the matrix supplied is always invertible.

In order to complete this assignment, you must do the following:

1. Fork the GitHub repository containing the stub R files at https://github.com/rdpeng/ProgrammingAssignment2 to create a copy under your own account.
2. Clone your forked GitHub repository to your computer so that you can edit the files locally on your own machine.
3. Edit the R file contained in the git repository and place your solution in that file (please do not rename the file).
4. Commit your completed R file into YOUR git repository and push your git branch to the GitHub repository under your account.
5. Submit to Coursera the URL to your GitHub repository that contains the completed R code for the assignment.

In addition to submitting the URL for your GitHub repository, you will need to submit the 40 character SHA-1 hash (as string of numbers from 0-9 and letters from a-f) that identifies the repository commit that contains the version of the files you want to submit. You can do this in GitHub by doing the following

1. Going to your GitHub repository web page for this assignment
2. Click on the “?? commits” link where ?? is the number of commits you have in the repository. For example, if you made a total of 10 commits to this repository, the link should say “10 commits”.
3. You will see a list of commits that you have made to this repository. The most recent commit is at the very top. If this represents the version of the files you want to submit, then just click the “copy to clipboard” button on the right hand side that should appear when you hover over the SHA-1 hash. Paste this SHA-1 hash into the course web site when you submit your assignment. If you don't want to use the most recent commit, then go down and find the commit you want and copy the SHA-1 hash.

A valid submission will look something like (this is just an example!)
```{r}
# https://github.com/rdpeng/ProgrammingAssignment2
```

```{r}
## Put comments here that give an overall description of what your
## functions do  

# The makeCacheMatrix() function takes a square invertible matrix as its  
# argument and creates a special list of four functions: set(), get(),  
# setsolve(), getsolve().   

# The cacheSolve() function takes a special list created by the makeCacheMatrix()  
# function as its argument and returns the inverse of a square matrix from the  
# special list.  


## Write a short comment describing this function
# The function below creates two objects, namely `x` and `inv` in the same  
# (enclosing) environment of the four functions: set(), get(), setsolve() and  
# getsolve(). By using the `<<-` operator, set() and setsolve() functions modify  
# the values of `x` and `inv` in the enclosing environment. The get() and  
# getsolve() simply retrieve the current values of `x` and `inv` in the enclosing  
# environment.

makeCacheMatrix <- function(x = matrix()) {
        inv <- NULL
        set <- function(y) {
                x <<- y
                inv <<- NULL
        }
        get <- function() x
        setsolve <- function(sol) inv <<- sol
        getsolve <- function() inv
        list(set = set, get = get, setsolve = setsolve, getsolve = getsolve)
}

## Write a short comment describing this function
# The function below checks if the value of `inv` have already been calculated  
# (exists) and, if exists, returns that `inv` value. If does not exist, it  
# retrieves the value of `x` in the enclosing environment, calculates the inverse  
# of it, modifies the value of `inv` in the enclosing environment, and returns  
# the value of the new `inv`. 

cacheSolve <- function(x, ...) {
        ## Return a matrix that is the inverse of 'x'
        inv <- x$getsolve()
        if(!is.null(inv)) {
                message("getting cached data")
                return(inv)
        }
        data <- x$get()
        inv <- solve(data, ...)
        x$setsolve(inv)
        inv
}
```

#### Let's try if the functions work
```{r}
m <- matrix(c(4, 2, 7, 6), nrow = 2, ncol = 2)
m

a <- makeCacheMatrix(m)
cacheSolve(a)
cacheSolve(a)
```

# Key Takeaway Functions

- **Loop Functions**  
  - lapply
    - *rnorm(n, mean = 0, sd = 1), runif(n, min = 0, max = 1)*
    
- **Debugging Tools**
  - Diagnosing the Problem
    - *invisible()*