---
title: "Week1_Material"
output: rmarkdown::github_document
---
## Background Material

Get working directory
```{r}
getwd()
```

Set a new working directory
```{r}
setwd("/Users/hyeonwooyang/Desktop/Desktop/5_Coursera/1_R_Programming/git_project/data")
```
```{r, setup, echo = FALSE}
knitr::opts_knit$set(root.dir = '/Users/hyeonwooyang/Desktop/Desktop/5_Coursera/1_R_Programming/git_project/data')
```

Display a list of files in the working directory
```{r}
dir()
```

Create a function
```{r}
myfunction <- function() {
        x <- rnorm(100)
        mean(x)
}
```



Execute a function
```{r}
myfunction()
```

Display a list of objects in the current environment
```{r}
ls()
```

Source(bring) all functions (objects) in a different file:  
`source("mycode.R")`

## Week 1: Getting Started and R Nuts and Bolts 

Assignment operator, `<-`
```{r}
x <- 5
msg <- "hello"
```

Display the numeric object (vector), `x`, that has one element
```{r}
print(x)        # explicit printing
x               # auto-printing
```
*The [1] indicated that x is a `vector` and 5 is the first element*

Display the character object (vector), `msg`, that has one element
```{r}
print(msg)
msg
```

Create integer sequences
```{r}
x <- 1:20
x
```

## Objects
Everything in R is an `object`. `Atomic` classes of objects:  

- character  
- numeric (real numbers)  
- integer  
- complex  
- logical (TRUE/FALSE)  



The most basic object is a `vector` 

- A vector can only contain objects of the same class  
BUT: The one exception is a "list", which is represented as a vector but can contain objects of different classes

## Numbers
Infinity  
`Inf`

Not a Number (undefined value)  
`NaN`

Attributes: `name`, `dimnames`, `dimensions` (e.g. matrices, arrays), `class`, `length`  
`attributes()`


## Vectors
Create vectors of objects `c()`
```{r}
x <- c(0.5, 0.6)        # numeric
x <- c(TRUE, FALSE)     # logical
x <- c(T, F)            # logical
x <- c("a", "b", "c")   # character
x <- 9:29               # integer
x <- c(1+0i, 2+4i)      # complex
```

`vector()` function
```{r}
x <- vector("numeric", length = 10)
x
```

## Mixing Objects        
Coerces to the least common denominator
```{r}
y <- c(1.7, "a")        # character
y <- c(TRUE, 2)         # numeric
y <- c("a", TRUE)       # character
```

Explicit Coercion: `as.`
```{r}
x <- 0:6
x

class(x)

as.numeric(x)
as.logical(x)
as.character(x)
```

Nonsensical coercion results in NAs
```{r}
x <- c("a", "b", "c")

as.numeric(x)
as.logical(x)
as.complex(x)
```

## List
```{r}
x <- list(1, "a", TRUE, 1 + 4i)
x
```
*[[1]] Elements of a list*  
*[1] 1 Elements of a vector*

## Matrices
Matrices are `vectors` with a `dimension attribute`  
Must have every element be the same class (object)
```{r}
m <- matrix(nrow = 2, ncol = 3)
m

dim(m)
attributes(m)
```


Matrices are constructed `column-wise`
```{r}
m <- matrix(1:6, nrow = 2, ncol = 3)
m
```

Matrices can also be created directly from `vectors` by adding a `dimension attribute`
```{r}
m <- 1:10
m

dim(m) <- c(2, 5)
m

```

Another way to create matrices by `cbind()` and `rbind()`
```{r}
x <- 1:3
y <- 10:12

cbind(x, y)
rbind(x, y)
```

## Factors
Factors are used to represent `categorical data`  
Factors can be `un-ordered` or `ordered`  
Factors are vectors with `labels`  
```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"))
x

table(x)
unclass(x)
```

Setting the order of the levels
```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"),
            levels = c("yes", "no"))
x

table(x)
unclass(x)
```

## Missing Values
`NA`  
`NaN`  
`is.na()` is used to test objects if they are NA  
`is.nan()` is used to test for NaN  
NA values have a class also, so there are integer NA, character NA, etc.  
A NaN value is also NA but the converse is not true (NaN means Not a Number, i.e. undefined value)  
```{r}
is.na(NaN)
is.nan(NA)
```

## Data Frames
Data frames are used to store tabular data  
Represented as a special type of list where every element of the list has to have the same length  
Each element of the list: column, and the length of each element of the list: the number of rows  
However, each column doesn't have to be the same type  

`row.names` - every row has its own name  
Data frames are created by calling `read.table()` or `read.csv()`  
Can be converted to a matrix by calling `data.matrix()` - force objects will be coerced to be the same class

```{r}
x <- data.frame(foo = 1:4, bar = c(T, T, F, F))
x

nrow(x)
ncol(x)
```

## Names Attribute  
Very useful for writing readable code and self-describing objects
```{r}
x <- 1:3
names(x)

names(x) <- c("foo", "bar", "norf") # name for each element
x
names(x)

x <- list(a = 1, b = 2, c = 3)
x

m <- matrix(1:4, nrow = 2, ncol = 2)
dimnames(m) <- list(c("a", "b"), c("c", "d"))
m
```

## Summary
Data Types:  

- atomic classes: numeric, logical, character, integer, complex
- vectors, lists
- factors
- missing values
- data frames
- names

## Reading Data  
Tabular Data: `read.table()`, `read.csv()` (returns a dataframe in R)  
Lines of a text file: `readLines()`  
R Code files: `source()`  
R Code files: `dget()` for reading R objects that have been dparsed into text files  
Reading binary objects into R: `load()` for reading in saved workspaces, `unserialize()` for reading single R objects in binary form  

## Writing Data
`write.table()`  
`writeLines()`  
`dump()`  
`dput()`  
`save()`  
`serialize()`  

`read.table()` function  
```{r}
str(read.table)
```

- `file`, the name of a file, or a connection  
- `header`, logical indicating if the file has a header line  
- `sep`, a string indicating how the columns are separated  
- `colClasses`, a character vector indicating the class of each column in the dataset  
- `nrows`, the number of rows in the dataset  
- `comment.char`, a character string indicating the comment character  
- `skip`, the number of lines to skip from the beginning  
- `stringAsFactors`, should character variables be coded as factors? `defualt = TRUE(yes)`  

e.g. `data <- read.table("foo.txt")`  
`read.csv` = `read.table` except that the default separator is a comma  

## Reading Large Tables
To prevent R from choking:  
Read the "help" page for `read.table()`, which contains many hints  
Make a rough calculation of the memory required to store your dataset  
Set `comment.char = ""` if there are no commented lines in your file  

the `colClasses` argument  
initial <- `read.table("datatable.txt", nrows = 100)`  
classes <- `sapply(initial, class)`   
tabAll <- `read.table("datatable.text", colClasses = classes)`  

## Textual Data Formats  
`dumping` and `dputing` preserves metadata(sacrificing some readability), so that another user doesn't have to specify it all over again.  
advantage: editable  
disadvanage: not space-efficient  

`dput()`
```{r}
y <- data.frame(a = 1, b = "a")
dput(y)

dput(y, file = "y.R")
new.y <- dget("y.R")
new.y
```

`dput()` function, essentially writes R code, which can be used to reconstruct an R object  

`dump()`  
`dget()` can only be used on a single R object  
`dump()` can be used on multiple R objects  
```{r}
x <- "foo"
y <- data.frame(a = 1, b = "a")

dump(c("x", "y"), file = "data.R")
rm(x, y) # remove x and y
source("data.R")

y
x
```

## Connections: Interfaces to the Outside World
`file` opens a connection to a file  
`gzfile` opens a connection to a file compressed with gzip  
`bzfile` opens a connection to a file compressed with bzip2  
`url` opens a connection to a webpage  

```{r}
str(file) # "str()" displays the internal structure of a R object

```

- `description` is the name of the file    
- `open` is a code indicating    
- `r` read only  
- `w` writing (and initializina new file)  
- `a` appending  
- `rb`, `wb`, `ab` reading, writing, or appending in binary mode (Windows)  

`con <- file("foo.txt", "r")`  
`data <- read.csv(con)`  
`close(con)`  

is the same as  
`data <- read.csv("foo.txt")`  

```{r}
connection <- url("http://www.jhsph.edu", "r")
x <- readLines(connection)
head(x)
```

## Subsetting: Basics  
`[` always returns an `object` of the same class as original (vector -> vector) (list -> list)  
Can be used to select more than one element of an object (there is one exception)  
`[[` extracts `elements` of a list or a dataframe  
Can only be used to extract a single element and the class of the returned object will not necessarily be a list or dataframe  
`$` extracts `elements` of a list or dataframe `by name`

```{r}
x <- c("a", "b", "c", "c", "d", "a")
x[1]
x[1:4]                          # numeric index
x[x > "a"]                      # logical index

u <- x > "a"
u
x[u]
```

## Subsetting: Lists
```{r}
x <- list(foo = 1:4, bar = 0.6)
x

x[1]

x[[1]]
x[[2]]

x$bar
x[["bar"]]      # = x[[2]]
x["bar"]        # = x[2]
```

```{r}
x <- list(foo = 1:4, bar = 0.6, baz = "hello")
x

x[c(1, 3)]

name <- "foo"
x[[name]]       # = x[[1]]
x$name          # element 'name' doesn't exist!
x$foo
```

```{r}
x <- list(a = list(10, 12, 14), b = c(3.14, 2.81))
x

x[[c(1, 3)]]
x[[1]][[3]]

x[[c(2, 1)]]
```

## Subsetting - Matrices
```{r}
x <- matrix(1:6, 2, 3)
x

x[1, 2]
x[2, 1]

x[1, ]
x[, 2]
```

By default, when a single element of a matrix is retrieved, it is returned as a `vector` of length 1 rather than a 1 x 1 matrix  
This behavior can be turned off by setting `drop = FALSE`  
```{r}
x[1, 2]
x[1, 2, drop = FALSE]

x[1, ]
x[1, , drop = FALSE]
```

## Subsetting - Partial Matching
```{r}
x <- list(aardvark = 1:5)
x

x$a
x[["a"]]
x [["a", exact = FALSE]]
```

## Subsetting - Removing Missing Values
```{r}
x <- c(1, 2, NA, 4, NA, 5)
is.na(x)
bad <- is.na(x)
x[!bad]
```

```{r}
x <- c(1, 2, NA, 4, NA, 5)
y <- c("a", "b", NA, "d", NA, NA)
complete.cases(x, y)

good <- complete.cases(x, y)
x[good]
y[good]
```

```{r}
airquality[1:6, ]
complete.cases(airquality[1:6, ])
good <- complete.cases(airquality)
airquality[good, ][1:6, ]
```

## Vectorized Operations
```{r}
x <- 1:4; y <- 6:9
x
y

x + y
x > 2
x >= 2
y == 8
x * y
x / y
```

```{r}
x <- matrix(1:4, 2, 2); y <- matrix(rep(10, 4), 2, 2)
x
y

x * y                   # element-wise multiplication
x / y
x %*% y                 # true matrix multiplication
```











<LINK>    
*Italic*  
**Bold**  

- bullet point  

Inserting R Code: `Cmd+Option+I`   
Avoid printing the code: `echo = FALSE`
